### 2018-06-04 Writing/SystematicBiology/Tab_pooledscharacters_test.tex
Cyst bill submission

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledscharacters_test}
\end{table}

>>>>

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscharacters_test}
\end{table}

### 2018-06-04 Writing/SystematicBiology/Tab_pooledsmethods_test.tex
Cyst bill submission

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:14 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledsmethods_test}
\end{table}

>>>>

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:14 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledsmethods_test}
\end{table}

### 2018-06-04 Writing/SystematicBiology/Tab_pooledstaxa_test.tex
Cyst bill submission

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledstaxa_test}
\end{table}

>>>>

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}

### 2018-06-04 Writing/SystematicBiology/draft.tex
Cyst bill submission

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

@@

\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise character differences for 100 characters. Blue colours correspond to low character differences and orange colours correspond to high character differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised (\textbf{C}), minimised (\textbf{D}) or randomise character difference (\textbf{E}). The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters. Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a low correlation between the characters.}
\label{Fig:modif_matrix}
\end{figure}

\subsection{Inferring topologies}
We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).

@@

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).]
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.

@@

Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g. when multiple trees are equal to the ``normal'' tree).
Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g. for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
\caption{\small{Effect of character difference on recovering the ``normal'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see Supplementary material 3 for the full summary statistics).
There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})

\input{Tables/Tab_pooledscenarios_test.tex}

@@

\input{Tables/Tab_pooledscharacters_test.tex}

@@

\input{Tables/Tab_pooledstaxa_test.tex}

@@

\input{Tables/Tab_pooledsmethods_test.tex}

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary materials @@@).

@@

The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.

@@

%TG: insisting on the shittiness of generatin correlation
Second, measuring and modifying character correlation is difficult. 
In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
%MB: We also don't really explore character congruence, that correlations can be ranked and spread over multiple internodes.
%MB: We didn't look at ranking and we didn't look at competing correlations. There's a lot more that could be done here so, we might talk about that.
%TG: not sure what you mean by the ranking stuff but it could definitely fit here.
However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e. some sets of characters can be highly correlated but effectively describe independent information).
Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.

@@

\subsection{Potential applications}
%TG: explaining that our "scenarios" are biased bootstraps
Effectively, our simulation protocol bootstraps our data ``with bias''.
In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.

%TG: explaining that we could use it for "clever" bootstrapping in big matrices
It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling (i.e. the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.

\subsection{Conclusion}
Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon to investigate phylogenetic relationships.
Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g. for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
%MB: I have no idea what thae above sentence means.
%TG: Modified. Is that clearer?
Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.

@@

% Suggested reviewers:
%- Liliana Davalos for MPE / Joe O'Reilly, April Wright

>>>>

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{outline.eps}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using maximum parsimony and Bayesian inference methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

@@

\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.

@@

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{Modif_matrix.eps}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise character differences for 100 characters. Blue colours correspond to low character differences and orange colours correspond to high character differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised (\textbf{C}), minimised (\textbf{D}) or randomise character difference (\textbf{E}). The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters. Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a low correlation between the characters.}
\label{Fig:modif_matrix}
\end{figure}

\subsection{Inferring topologies}
We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).

@@

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.

@@

Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}). 
Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g. when multiple trees are equal to the ``normal'' tree).
Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g. for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).

@@

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{RF_results_best.eps}
\caption{\small{Effect of character difference on recovering the ``normal'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{Tr_results_best.eps}
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see supplementary material 3 for the full summary statistics).
There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & maxi:mini & 0.573 & 356436.000 & \textbf{0} \\ 
   & maxi:rand & 0.873 & 287225.000 & \textbf{0} \\ 
   & mini:rand & 0.856 & 95841.500 & \textbf{0} \\ 
  Tr & maxi:mini & 0.614 & 358800.000 & \textbf{0} \\ 
   & maxi:rand & 0.908 & 288223.500 & \textbf{0} \\ 
   & mini:rand & 0.858 & 98507.500 & \textbf{0} \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled scenarios. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscenarios_test}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscharacters_test}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledsmethods_test}
\end{table}

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).

@@

The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.b

@@

Second, measuring and modifying character correlation is difficult. 
In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e. some sets of characters can be highly correlated but effectively describe independent information).
Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.

@@

\subsection{Potential applications}
Effectively, our simulation protocol bootstraps our data ``with bias''.
In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.

It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling (i.e. the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.

\subsection{Conclusion}
Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon to investigate phylogenetic relationships.
Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g. for characters hierarchy][]{DeLaet2015,BrazeauNA} already exist.
However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.

@@
